---
title: "Next Word Prediction"
author: "Wanjie Feng"
date: "2024-04-15"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Project Overview

Around the world, people are spending an increasing amount of time on their mobile devices for email, social networking, banking and a whole range of other activities. But typing on mobile devices can be a serious pain. SwiftKey, our corporate partner in this capstone, builds a smart keyboard that makes it easier for people to type on their mobile devices. One cornerstone of their smart keyboard is predictive text models. When someone types:

*I went to the*

the keyboard presents three options for what the next word might be. For example, the three words might be *gym, store, restaurant*. 

In this capstone, the text data logged from News, Twitter, and Blogs are analyzed and processed. These data are used to build a model to predict the next word when giving a phrase. 



## Outline of the Project

- Data download and read in 
- Exploratory Data Analysis
- Build Katz backoff model
- Predict Next Words
- Shiny App


## Exploratory Data Analysis

```{r, echo = FALSE, warning=FALSE, message=FALSE}
library(wordcloud)
library(RColorBrewer)
gram_1 <- read.csv("gram_1.csv")
wordcloud(gram_1$Word, gram_1$Counts, scale=c(3.5, 0.02), max.words=200, random.order=FALSE,
          rot.per=0.35, colors=brewer.pal(5, "Dark2"))
```

## Katz Backoff Model

```{r, echo = FALSE, warning=FALSE, message=FALSE}
library(patchwork)
library(magick)
(image_read("katz.svg")|>image_ggplot())
```

## Predict Next Word and Shiny App

- The R markdown and Shiny App files can be found at https://github.com/Aaron2014/JHU_Capstone_Project

- The Shiny App can be accessed at https://es0for-aaron-fong.shinyapps.io/predict_word/
